---
title: "eda"
output: github_document
---

This is for visualization with ggplot2 project!

```{r, echo=FALSE}
library(tidyverse)
```

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = case_match(
      id, 
      "USW00094728" ~ "CentralPark_NY", 
      "USW00022534" ~ "Molokai_HI",
      "USS0023B17S" ~ "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = floor_date(date, unit = "month")) |>
  select(name, id, everything())
```


Let's make some plots

```{r}
 weather_df |> 
  ggplot(aes(x = prcp)) +
  geom_histogram()
```

first, find the extram number in the histogram to see if it is a normal number
or an outlier
```{r}
 weather_df |>
   filter(prcp > 1000)
```

next, zoom in part of the data
```{r}
 weather_df |>
  filter(tmax > 20, tmax < 30) |> 
  ggplot(aes(x = tmin, y = tmax, color = name, shape = name)) +
  geom_point()
```


## group_by

```{r}
 weather_df |>
  group_by(name)
```

counting stuff

```{r}
 weather_df |>
  group_by(name) |> 
  summarize(n_obs = n())
```

```{r}
 weather_df |>
  group_by(name) |> 
  summarize(
    n_obs = n(),
    n_dist = n_distinct(month))
```


```{r}
 weather_df |>
  group_by(name, month) |> 
  summarize(
    n_obs = n()
    )
```

can specifically use count function
```{r}
 weather_df |>
  count(name)
```


## 2x2

```{r}
 weather_df |> 
  drop_na(tmax) |> 
  filter(name != "Molokai") |> 
  mutate(
    cold = case_when(
      tmax < 5 ~ "cold",
      tmax >= 5 ~ "not cold"
    )
  ) |> 
  group_by(name, cold) |> 
  summarise(count = n())

 weather_df |> 
  drop_na(tmax) |> 
  filter(name != "Molokai") |> 
  mutate(
    cold = case_when(
      tmax < 5 ~ "cold",
      tmax >= 5 ~ "not cold"
    )
  ) |> 
   janitor::tabyl(name,cold) # another function creating table
```


## general numeric summaries.

let's try some other usefual summaries.

```{r}
 weather_df |> 
  group_by(name, month) |> 
  summarise(
    mean_tmax = mean(tmax, na.rm = TRUE),
    median_tmin = median(tmin, na.rm= TRUE),
    sd_prcp = sd(prcp, na.rm = TRUE)
  )
```


summarize and then plot

```{r}
 weather_df |> 
  group_by(name, month) |> 
  summarise(
    mean_tmax = mean(tmax, na.rm = TRUE),
    median_tmin = median(tmin, na.rm= TRUE),
    sd_prcp = sd(prcp, na.rm = TRUE)
  ) |> 
  ggplot(aes(x= month, y = mean_tmax, color = name)) +
  geom_point() +
  geom_line()
```


format for readers

```{r}
 weather_df |> 
  group_by(name, month) |> 
  summarise(
    mean_tmax = mean(tmax, na.rm = TRUE),
  ) |> 
  pivot_wider(
    names_from = name,
    values_from = mean_tmax
  ) |> 
  knitr::kable(
    digit = 3, # how many digit want to save
    col.names = c("Month", "Central Park", "Molokai", "Waterhole"))
```


## grouped mutates

```{r}
 weather_df |> 
  group_by(name) |> 
  mutate(mean_tmax = mean(tmax, na.rm = TRUE),
         centered_tmax = tmax - mean_tmax) |> 
  ggplot(aes(x = date, y = centered_tmax, color = name)) +
  geom_point()
```



Find hottest / coldest days


```{r}
 weather_df |> 
  group_by(name) |> 
  mutate(
    temp_rank = min_rank(desc(tmax))
  ) |> 
  filter(temp_rank < 4)

 weather_df |> 
  group_by(name) |>
  filter(min_rank(tmax) < 4) |> 
  arrange(tmax)
```


```{r}
 weather_df |> 
  group_by(name) |> 
  mutate(
    lagged_tmax = lag(tmax),
    temp_change = tmax - lagged_tmax
  ) |> 
  filter(min_rank(temp_change) < 3)
```


```{r}
 weather_df |> 
  group_by(name) |> 
  mutate(
    lagged_tmax = lag(tmax),
    temp_change = tmax - lagged_tmax
  ) |> 
 summarize(
   sd_tmax_change = sd(temp_change, na.rm = TRUE)
 )
```

Learning assessment










