---
title: "viz_and eda"
author: "cathy"
date: "2024-10-01"
output: github_document
---

This is for visualization with ggplot2 project!

```{r}
library(tidyverse)
library(patchwork)
library(ggplot2)
library(haven)
```

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = case_match(
      id, 
      "USW00094728" ~ "CentralPark_NY", 
      "USW00022534" ~ "Molokai_HI",
      "USS0023B17S" ~ "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```


Make a scartterplot but fancy this time.

```{r}
 weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3) +
  labs(
    title = "Temperature sctterplot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    color = "Location",
    caption = "Weather data taken from rnoaa package for three stations"
  )
```


Scales -- star with 'x' and 'y'

```{r}
 weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3) +
  labs(
    title = "Temperature sctterplot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    color = "Location",
    caption = "Weather data taken from rnoaa package for three stations"
  ) +
  scale_x_continuous(
    breaks = c (-15, 0, 20),
    labels = c("-15C", "0", "20")
  ) +
  scale_y_continuous(
    limits = c (0,30),
    transform = "sqrt"
  )
```


Look at color:

```{r}
ggp_scatterplot = 
 weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3) +
  labs(
    title = "Temperature sctterplot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    color = "Location",
    caption = "Weather data taken from rnoaa package for three stations"
      ) +
  viridis::scale_color_viridis(discrete = TRUE)

ggp_scatterplot 
```

also can use " scale_color_hue(h = c(100, 400)" to change the number, change the color


## Themes

```{r}
ggp_scatterplot +
  theme(legend.position = "bottom")
```


change the background black and white
have order matters

show different types of the plot virsion

```{r}
ggp_scatterplot +
  theme_bw() +
  theme(legend.position = "bottom")
```

```{r}
ggp_scatterplot +
  theme(legend.position = "bottom") +
  theme_bw()
```

```{r}
ggp_scatterplot +
  theme(legend.position = "bottom") +
  theme_minimal()
```


```{r}
ggp_scatterplot +
  theme(legend.position = "bottom") +
  theme_classic()
```


Learning assessment
```{r}
 weather_df |> 
  ggplot(aes(x= date, y =tmax, color = name, size = prcp)) +
  geom_point(alpha = .3) +
  geom_smooth(se = FALSE) +
  viridis::scale_color_viridis(discrete = TRUE) + 
  labs(
    x = "Date",
    y = "Maxiumum daily temperature (C)",
    title = "Seasonal variation in Max Temp",
    size = "Precipitation"
 ) +
  theme_minimal() + 
  theme(legend.position = "bottom")
```


Extra bonus stuff in 'ggplot'

Use different datasets in differernt 'gemo's

```{r}
central_park_df =
  weather_df |> 
  filter(name == "CentralPark_NY")

molokai_df =
  weather_df |> 
  filter(name == "Molokai_HI")

 weather_df |> 
  ggplot(aes(x = date, y =tmax, color= name)) +
  geom_point() +
  geom_line(data = central_park_df)
```

## Multiple panels

good to seperate by names
```{r}
weather_df |> 
  ggplot(aes(x = tmax, fill =name)) +
  geom_density() +
  facet_grid(. ~ name)
```


put different graph together
```{r}
ggp_tmax_tmin = 
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color =name)) +
  geom_point(alpha = .3)


ggp_tmax_density = 
  weather_df |> 
  ggplot(aes(x = tmax, fill = name)) +
  geom_density(alpha = .3) 

ggp_tmax_date =
  weather_df |> 
  ggplot(aes(x = date, y = tmax, color =name)) +
  geom_point() +
  geom_smooth(se = FALSE)

  (ggp_tmax_tmin + ggp_tmax_density)/ ggp_tmax_date 
```

try to remove every theme, but only remove one
ggp_together =
  (ggp_tmax_tmin + ggp_tmax_density)/ ggp_tmax_date +
  theme(legend.position = "none")



## Data manipulation

```{r}
weather_df |> 
  ggplot(aes(x = name, y = tmax, fill =name)) +
  geom_violin(alpha =.5)
```

```{r}
weather_df |> 
  mutate(name = fct_relevel(name, c("Molokia_HI", "CentralPark_NY", "Waterhole_WA"))) |> 
  ggplot(aes(x = name, y = tmax, fill =name)) +     
  geom_violin(alpha =.5)
```



PULSE data next.

```{r}
pulse_df =
  read_sas("data/public_pulse_data.sas7bdat") |> 
  janitor::clean_names() |> 
  pivot_longer(
    cols = bdi_score_bl:bdi_score_12m,
    names_to = "visit",
    values_to = "bdi_score",
    names_prefix = "bdi_score"
  ) |> 
  mutate(visit = ifelse(visit == "bl", "00m", visit))

pulse_df|> 
  ggplot(aes(x = visit, y = bdi_score)) +
  geom_boxplot()
```




