---
title: "viz_and eda"
author: "cathy"
date: "2024-09-26"
output: github_document
---

This is for visualization with ggplot2 project!

```{r}
library(tidyverse)
library(ggridges)
```

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = case_match(
      id, 
      "USW00094728" ~ "CentralPark_NY", 
      "USW00022534" ~ "Molokai_HI",
      "USS0023B17S" ~ "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

Making our first plot :-)

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax))
```

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
```

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```

```{r}
 ggp_weather_scatterplot = 
   weather_df |>
   ggplot(aes(x = tmin, y = tmax)) + 
   geom_point()
  
ggp_weather_scatterplot
```


Check why some rows are missing
```{r}
 weather_df |> 
  filter(is.na(tmax))
```

## Fancier scatterplot
```{r}
   weather_df |>
   ggplot(aes(x = tmin, y = tmax, color = name)) + 
   geom_point(alpha = 0.3, size = 0.8)+
   geom_smooth(se = FALSE)
```

Where you difine aesthetics can matter

here, geom_point follow the command "only put name in color"
but geom_smooth is only follow ggpolt, don't know anything about the color, follow the top line of ggplot
```{r}
   weather_df |>
   ggplot(aes(x = tmin, y = tmax)) + 
   geom_point(aes(color = name), alpha = 0.3, size = 0.8) +
   geom_smooth(se = FALSE)
```

use faceting real quick
 
```{r}
   weather_df |>
   ggplot(aes(x = tmin, y = tmax, color = name)) + 
   geom_point(alpha = .3) +
   geom_smooth(se = FALSE)+
  facet_grid(. ~ name)
```

```{r}
   weather_df |>
   ggplot(aes(x = tmin, y = tmax, color = name)) + 
   geom_point(alpha = .3) +
   geom_smooth(se = FALSE)+
   facet_grid(name ~ .)
```


Let's make a somewhat more interesting scatterplot

```{r}
 weather_df |>
   ggplot(aes(x = date, y = tmax, color = name, size = prcp)) +
   geom_point(alpha = .3) +
   geom_smooth(se = FALSE) +
   facet_grid(. ~ name)
```

Learning assessment

```{r}
    weather_df |>
    filter (name == "CentralPark_NY") |> 
    mutate(
      tmax_fahr = tmax * (9/5)+32,
      tmin_fahr = tmin * (9/5)+32
    ) |> 
    ggplot(aes(x = tmin_fahr, y = tmax_fahr, color = name)) + 
    geom_point(alpha = .3) +
    geom_smooth(method = "lm", se = FALSE)
```


## Small things
if only want the line, can just draw the line, not put anything else into code
```{r}
 weather_df |>
   ggplot(aes(x = tmin, y = tmax)) +
   geom_smooth(se = FALSE)
```

```{r}
 weather_df |>
   ggplot(aes(x = tmin, y = tmax)) +
   geom_hex()
```

if it's a variable then do not need quate, if it's doesn't a variable name, need quate to let R know
```{r}
 weather_df |>
   ggplot(aes(x = tmin, y = tmax)) +
   geom_point(color ="blue")
```


here, ggplot identify "blue" as a variable, it acutally create a variable and name it "blue"
```{r}
 weather_df |>
   ggplot(aes(x = tmin, y = tmax, color ="blue")) +
   geom_point()
```


## Univarite plots

```{r}
 weather_df |> 
 ggplot(aes(x = tmin)) + 
   geom_histogram()
```

```{r}
 weather_df |> 
 ggplot(aes(x = tmin, color = name)) + 
   geom_histogram()
```

```{r}
 weather_df |> 
 ggplot(aes(x = tmin, fill = name)) + 
   geom_histogram()
```

```{r}
 weather_df |> 
 ggplot(aes(x = tmin, fill = name)) + 
   geom_histogram(position = "dodge")
```

how would I fix this? maybe facet?

```{r}
 weather_df |>
  ggplot(aes(x = tmax, fill = name)) + 
  geom_histogram () +
  facet_grid(. ~ name)
```

mayber a density plot?
```{r}
 weather_df |>
  ggplot(aes(x = tmax, fill = name)) + 
  geom_density ()
```


```{r}
 weather_df |>
  ggplot(aes(x = tmax, fill = name)) + 
  geom_density (alpha = .3)
```


maybe a box plot?
```{r}
 weather_df |>
  ggplot(aes(x = name, y = tmin, fill = name)) + 
  geom_boxplot ()
```

violin plots
```{r}
 weather_df |>
  ggplot(aes(x = name, y = tmin, fill = name)) + 
  geom_violin ()
```

ridge plot
```{r}
 weather_df |>
  ggplot(aes(x = tmin, y = name)) + 
  geom_density_ridges()
```


LA univarite plots

```{r}
weather_df |>
  ggplot(aes(x = prcp, fill = name)) + 
  geom_histogram (position = "dodge")

 weather_df |>
  ggplot(aes(x = prcp, fill = name)) + 
  geom_density (alpha = .3)

weather_df |>
  ggplot(aes(x = name, y = prcp)) + 
  geom_boxplot ()

weather_df |>
  ggplot(aes(x = name, y = prcp, fill = name)) + 
  geom_violin ()

weather_df |>
  filter(prcp > 10, prcp < 1000) |>
  ggplot(aes(x = prcp, fill = name)) + 
  geom_density (alpha = .3)

weather_df |>
  ggplot(aes(x = prcp, y = name)) + 
  geom_density_ridges(scale = .85)

```

## Saving and embedding plots

```{r}
ggp_weather = 
weather_df |>
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point ()

ggsave("plots/ggp_weather.pdf", ggp_weather, width = 8, height = 6)
```

embedding plots

```{r, fig.width = 6, fig.asp = .6}
ggp_weather = 
weather_df |>
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point ()

```




